Procedural generation of levels for the angry birds videogame using evolutionary computation

### 2019-02-28 Springer_IntlCongress/author/authorsample.tex
Modification to the book chapter document

\title*{Procedural generation of levels for the angry birds videogame using evolutionary computation}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\author{\textbf{Jaime Salinas-Hernández and Mario Garcia-Valdez}}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{Jaime Salinas-Hernández \at Tijuana Institute of Technology, Tijuana, México \email{salinas.jaime1217@gmail.com}
\and Mario Garcia-Valdez(\Letter) \at Tijuana Institute of Technology, Tijuana, México \email{mariosky@gmail.com}}
%
% You can put both on the same level as in:
% Name
% e-mail
% Name 2
% e-mail 2
% Institution
%
% Use the package "url.sty" to avoid
% problems with special characters
% used in your e-mail or web address
%
\titlerunning{Procedural generation of levels for the angry birds videogame}
\maketitle

@@

\section{Introduction}
\label{sec:1}
Procedural content generation is an interesting subject on the entertainment area, primarily in the video game industry because of the easines and quickness provided to generate content for diferent video game projects, in this project a evolutionary computation-based system is proposed to generate complex content for videogames based on an evolutionary algorithm and open-ended- evolution.

For this project a procedural generation content system is proposed for the game angly birds, angry birds is a game developed by Rovio Entrtainment in which the objective is to destroy structures placed on the map as well as eliminating green pig-like entities in order to obtain the highest score possible, the game mechanics is to use a slingshot to shoot a certain number of birds with different abilities to destroy the structures and the objectives. \cite{RovioEntertainmentCorporation2009}

While it is possible to create a evolutionary agent to play the levels and obtain the best score the main objective of this project is to generate the levels can ca be played by the agents or physical players, the levels themselves need to be complex in the sense that interesting structures can be created but it has to be possible to complete in some way.

In Section 2, a background of the subjects used on this project will be presented, this subjects, while not all being used on by different authors on their works provide a great help on the development of the proposed system, Section 3, contains a quick overview of some aspects used by different authors in order to tackle this problem, their design and basic functionality, in Section 4 the description of the current state of the project is explained, from the basis of the development to the diferent menthods that have been used to progress on the different sections of the project, in Section 5 we explain the future work of the project, where we are in the development phase and how the future implementations are to further improve the project and finally, in Section 6 we present our conclusions

\section{Background}
\label{sec:3}
In this Section we provide a quick overview of the literature about the areas that provide the basis to the development of the current project, this section is divided in three parts, the fisrt one gives us an explanation of what procedural content generation entails, the second one provides an explanation on genetic algorithms and how can they be used to solve optimization problems, and finally, the third one provides us with an overview of Open-Ended Evolution and what it means to have an Open-Ended Evolutionary algorithm.

@@

M. Kaidan et al. \cite{Kaidan2015} propose a model in which the generation of content is controlled by the skill of a player while actively playing the game, the moment the player finishes a certain level the algorithm obtains the score and generates the next level to be played. The generation of the levels is based on a formula which take into account the average velocities to evaluate the levels however the generation of the levels did not take into account the number of birds and pigs placed on the level and as such the levels wee able to be clared with ease by the participants.

@@

Y. Jiang et al. \cite{Jiang2017} proposed the use of predefined letter style patterns and combined to create a small pool of words used to be combined in order to create levels with text on them, the participants would play the game and after that they would be asked if they liked the layout of the levels, the way the requirements for the competition were fullfilled were by creating the letter structures as stable as possible, use the provided layered structures, generating a said number of usable birds and pigs by using an heuristic calculation.

@@

In this Section we describe the approach that was used for this project, 
The game contains a total of 11 pieces as shown on figure 2 that cannot be modified, it is not possible to add more or modify the existing ones, with this in mind the purpose is to create bundles of this same pieces in different locations and angles in order to create structures that can be used as building blocks for more complex levels, in order to do this a sequence of events needs to be done as follows:

@@

The way to represent the elements of a member of the population is by using the id of each piece and the new added ones has a pointer to the respective data that needs to be obtained, this way and individual can be represented as shown on figure 5 where each chromosome represents a certain item in the pool of pieces, the numbers for this elements are assigned in a first come first served basis, the numbers are infinitely incremented and this same pieces can be repeated on each individual of the population

@@

The evaluation process for this project is still in the development stages, however in order to check that the evolution is working has it should a criteria has been temporarily made as follows, each individual in the population has 100 points so before the simulation starts each of them are viable candidates to be a elite member of the population regardless relative height or complexity, then the remaining of the process is separated in three phases has follows.

@@

\section{Future work}
Since there is a need to create complex structures than can cover a great section of the play map, a method to evaluate the distribution is needed, for this instance we propose using the rule of thirds \cite{DarrenRowse} which is a guideline used while creating imagen in which the purpose is to create aesthetical pleasant images where the focus point or the most important area is at the center of the image as shown on figure 7, using this rule different distribution groups or masks are created has shown on figure 8 in order to use them before the simulation of an individual takes place in order to distribute the pieces in a way that can cover more area and create more balanced levels instead of using a tower like distribution as shown on figure 9.
Using this same distribution a new evaluation will be added in which according to how much of the area is filled a score will be given and this score will be used as a new sorting value at the moment of selecting individuals for the crossover operation, since as shown on figure 8 there is a possibility that a huge mask is given to an individual it can be possible that the amount of pieces will not be able to cover the full area given, so in order to prevent this kind of problems we also propose to give each mask a value that refers to the minimum amount of pieces and height required to fill the most of the mask then check the amount of pieces and height before simulation for a single individual and then randomize which mask of the ones that can be used will be assigned to that individual

@@

\begin{figure}[htbp]
    \centerline{\includegraphics[width=80mm]{Images/mask_distribution.png}}
    \caption{Rule of thirds applied to an individual.}
    \label{fig}
\end{figure}

@@

The use of artificial intelligence in videogames is a common thing almost since before the first generation of videogames on 1972 as simply entities that would follow a set group of actions but as the generation kept advancing the need for more intelligent entities appeared, as of this generation some videogames use AI as ways to improve the interaction between man and game, but also using it in order to facilitate the way videogames are created, so for this project the focus of using evolutionary algorithms as great interest, as a starting point a rather simple game can be used as a workbench of the work but it also helps us define how much more can be done when all the required objects are obtained, improve on the same system and find ways for it to be used on different other games

>>>>

\title*{Procedural generation of levels for the angry birds videogame using evolutionary computation}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\author{\textbf{Jaime Salinas-Hernández and Mario Garcia-Valdez}}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{Jaime Salinas-Hernández \at Tijuana Institute of Technology, Tijuana, México \email{salinas.jaime1217@gmail.com}
\and Mario Garcia-Valdez(\Letter) \at Tijuana Institute of Technology, Tijuana, México \email{mariosky@gmail.com}}
%
% You can put both on the same level as in:
% Name
% e-mail
% Name 2
% e-mail 2
% Institution
%
% Use the package "url.sty" to avoid
% problems with special characters
% used in your e-mail or web address
%
\titlerunning{Procedural generation of levels for the angry birds videogame ...}
\maketitle

@@

\section{Introduction}
\label{sec:1}
Procedural content generation is an interesting subject on the entertainment area, primarily in the video game industry because of the easiness and quickness provided to generate content for different video game projects, in this project a evolutionary computation-based system is proposed to generate complex content for videogames based on an evolutionary algorithm and open-ended- evolution.

For this project a procedural generation content system is proposed for the game angry birds, angry birds is a game developed by Rovio Entertainment in which the objective is to destroy structures placed on the map as well as eliminating green pig-like entities in order to obtain the highest score possible, the game mechanics is to use a slingshot to shoot a certain number of birds with different abilities to destroy the structures and the objectives. \cite{RovioEntertainmentCorporation2009}

While it is possible to create an evolutionary agent to play the levels and obtain the best score the main objective of this project is to generate the levels can ca be played by the agents or physical players, the levels themselves need to be complex in the sense that interesting structures can be created but it has to be possible to complete in some way.

In Section 2, a background of the subjects used on this project will be presented, this subjects, while not all being used on by different authors on their works provide a great help on the development of the proposed system, Section 3, contains a quick overview of some aspects used by different authors in order to tackle this problem, their design and basic functionality, in Section 4 the description of the current state of the project is explained, from the basis of the development to the different methods that have been used to progress on the different sections of the project, in Section 5 we explain the future work of the project, where we are in the development phase and how the future implementations are to further improve the project and finally, in Section 6 we present our conclusions.

\section{Background}
\label{sec:3}
In this Section we provide a quick overview of the literature about the areas that provide the basis to the development of the current project, this section is divided in three parts, the first one gives us an explanation of what procedural content generation entails, the second one provides an explanation on genetic algorithms and how can they be used to solve optimization problems, and finally, the third one provides us with an overview of Open-Ended Evolution and what it means to have an Open-Ended Evolutionary algorithm.

@@

M. Kaidan et al. \cite{Kaidan2015} propose a model in which the generation of content is controlled by the skill of a player while actively playing the game, the moment the player finishes a certain level the algorithm obtains the score and generates the next level to be played. The generation of the levels is based on a formula which take into account the average velocities to evaluate the levels however the generation of the levels did not take into account the number of birds and pigs placed on the level and as such the levels were able to be cleared with ease by the participants.

@@

Y. Jiang et al. \cite{Jiang2017} proposed the use of predefined letter style patterns and combined to create a small pool of words used to be combined in order to create levels with text on them, the participants would play the game and after that they would be asked if they liked the layout of the levels, the way the requirements for the competition were fulfilled were by creating the letter structures as stable as possible, use the provided layered structures, generating a said number of usable birds and pigs by using an heuristic calculation.

@@

In this Section we describe the approach that was used for the development of this project, first presenting a quick overview of the integration of a genetic algorithm with the combination of pieces, the way the open-ended segment of the project handled the creation of new compounds for the pool of pieces and then an explanation on the way the individuals are evaluated.

The game contains a total of 11 pieces as shown on figure 2 that cannot be modified, it is not possible to add more or modify the existing ones, with this in mind the purpose is to create bundles of this same pieces in different locations and angles in order to create structures that can be used as building blocks for more complex levels, in order to do this a sequence of events needs to be done as follows:

@@

The way to represent the elements of a member of the population is by using the id of each piece and the new added ones has a pointer to the respective data that needs to be obtained, this way and individual can be represented as shown on figure 5 where each chromosome represents a certain item in the pool of pieces, the numbers for this elements are assigned in a first come first served basis, the numbers are infinitely incremented and this same pieces can be repeated on each individual of the population.

@@

The evaluation process for this project is still in the development stages, however in order to check that the evolution is working has it should a criteria has been temporarily made as follows, each individual in the population has 100 points so before the simulation starts each of them are viable candidates to be an elite member of the population regardless relative height or complexity, then the remaining of the process is separated in three phases has follows.

@@

Using the proposed approach a modification to the generation of individuals was developed, this modification allowed the generation to create structures that are able to maintain balance except for some cases where some element of the individual where irregular pieces, this pieces like triangles and circles when placed in positions between the base of the tower to before the top caused the tower to fall depending if the subsequent pieces in the individual where able to outbalance the structure.

\begin{figure}[htbp]
    \centerline{\includegraphics[width=80mm]{Images/results_v1.png}}
    \caption{Results with current approach.}
    \label{fig}
    \end{figure}

With this in consideration the results as shown in figure 6 and subsequently figure 7 where able to be obtained, however since the objective of the project is to construct more complex structures the obtained results will be used as a base for the further development of the project with the implementation of the concepts presented in the future work section.

\section{Future work}
Since there is a need to create complex structures than can cover a great section of the play map, a method to evaluate the distribution is needed, for this instance we propose using the rule of thirds \cite{DarrenRowse} which is a guideline used while creating imagen in which the purpose is to create aesthetical pleasant images where the focus point or the most important area is at the center of the image as shown on figure 8, using this rule different distribution groups or masks are created has shown on figure 9 in order to use them before the simulation of an individual takes place in order to distribute the pieces in a way that can cover more area and create more balanced levels instead of using a tower like distribution as shown on figure 10.

Using this same distribution a new evaluation will be added in which according to how much of the area is filled a score will be given and this score will be used as a new sorting value at the moment of selecting individuals for the crossover operation, since as shown on figure 8 there is a possibility that a huge mask is given to an individual it can be possible that the amount of pieces will not be able to cover the full area given, so in order to prevent this kind of problems we also propose to give each mask a value that refers to the minimum amount of pieces and height required to fill the most of the mask then check the amount of pieces and height before simulation for a single individual and then randomize which mask of the ones that can be used will be assigned to that individual.

@@

\begin{figure}[htbp]
    \centerline{\includegraphics[width=80mm]{Images/mask_distribution.png}}
    \caption{Masks examples with the rule of thirds.}
    \label{fig}
\end{figure}

\begin{figure}[htbp]
    \centerline{\includegraphics[width=80mm]{Images/chromosome_thirds.png}}
    \caption{Rule of thirds applied to an individual.}
    \label{fig}
\end{figure}

Another proposed method for the evolution of the individuals is to use a mutation operation that is capable of adding or removing pieces from an individual, this way the diversity of the population will be greater on future generations, also by using the same mutation idea a change on the mask assigned to a particular individual can be modified as well, in the case a mask is either too small or too big for an individual because of the orientation of the pieces a precaution has been created, in case the mask is too small for the number of pieces after filling the mask areas the code will take the remaining pieces and begin placing them as if the mask was repeated, and in the case the number of pieces is too small the system will simply return the pieces that were able to be fitted in and continue with the process as usual.

@@

The use of artificial intelligence in videogames is a common thing almost since before the first generation of videogames on 1972 as simply entities that would follow a set group of actions but as the generation kept advancing the need for more intelligent entities appeared, as of this generation some videogames use AI as ways to improve the interaction between man and game, but also using it in order to facilitate the way videogames are created, so for this project the focus of using evolutionary algorithms as great interest, as a starting point a rather simple game can be used as a workbench of the work but it also helps us define how much more can be done when all the required objects are obtained, improve on the same system and find ways for it to be used on different other games.
